Linear Stochastic Approximation- Constant Step-Size and Iterate Averaging

### 2017-05-11 instance.tex
revisions

\section{Main Results}\label{sec:mainresults}
In this section, we derive instance dependent bounds  which is valid for a fixed $P$ and in the \Cref{sec:uniform}, address the question of deriving uniform bounds $\forall\,P\in \P$, where $\P$ is a class of distributions over $\C^{\dcd}$. We first present the results and then follow it up with a discussion.
\begin{lemma}\label{lm:pd}
Let $P$ be a positive definite distribution over $\C^{\dcd}$, then there exists an $\alpha_P$ such that $\rhod{P}>0$ and $\rhos{P}>0,~\forall \alpha \in (0,\alpha_P)$.
\end{lemma}
\begin{comment}
\begin{proof}
\begin{align*}
\rhos{P}&=\inf_{x:\norm{x}=1}x^* (A_P+A_P^*)x -\alpha x^*\EE{A_t^* A_t} x\\
&=\inf_{x:\norm{x}=1}x^* (A_P+A_P^*)x -\alpha x^* A^*_P A_P -\alpha x^* \EE{M_t^* M_t} x\\
&\geq \lambda_{\min}(A^*_P+A_P)-\alpha \norm{A_P}^2-\sigma^2_P
\end{align*}
The proof is complete by choosing $\alpha_P<\frac{\lambda_{\min}(A^*_P+A_P)}{\norm{A_P}^2+\sigma^2_P}$
\end{proof}
\end{comment}
\begin{lemma}\label{lm:simtran}[Similarity Transformation]
Let $P^{AS}$ be asymptotically stable, then there exists a $U\in \gln$ such that $P^{AS}_U$ is positive definite.
\end{lemma}

\begin{theorem}\label{th:pdrate}
Let $P$ in \Cref{assmp:lsa}-~\ref{dist} be a positive definite distribution over $\C^{\dcd}$. Then
\begin{align*}
\EE{\norm{\thh_t-\ts}^2}
\leq
\left(1+\frac2{\alpha\rhod{P}}\right)\frac{1}{\alpha \rhos{P}}
\left(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} \right)\,.
%\left(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\theta_0-\ts})}{t+1} \right)\,.
\end{align*}

\end{theorem}
\begin{theorem}\label{thm:simtran}[Change of Basis]
Let $P$ in \Cref{assmp:lsa}-~\ref{dist} be AS. Define $\gamma_t\eqdef U^{-1}\theta_t,\,,\gamma^*\eqdef U^{-1}\ts$, and suppose if $P_U$ is positive definite, then
\begin{align*}
\EE{\norm{\gh_t-\gamma^*}^2}
\leq
\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\norm{U^{-1}}^2}{\alpha \rhos{P_U}}
\left(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} \right)\,.
%\left(\frac{\norm{\gamma_0-\gamma^*}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\gamma_0-\gamma^*}}{t+1} \right)\,.
%\left(\frac{\norm{\gamma_0-\gamma^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\gamma_0-\gamma^*})}{t+1} \right)\,,
\end{align*}
where $\gh_t=\frac{1}{t+1}\sum_{s=0}^t \gamma_s$.
\end{theorem}
\begin{corollary}\label{cor:as}
Let $P$ in \Cref{lsadist} be asymptotically stable. Then, there exists
\begin{align*}
\EE{\norm{\thh_t-\ts}^2}
\leq
\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\cond{U}^2}{\alpha \rhos{P_U}}
\left(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} \right)\,.
\end{align*}
\end{corollary}
\begin{theorem}[Lower Bound]
There exists a $P$ such that
\begin{align}
\frac{1}{(t+1)^2}(\alpha \lambda_{\min})^{-2}(1-\alpha \lambda_{\min}^t) \theta^2_0(1)\leq \norm{\eh_t}^2\leq
\frac{1}{(t+1)^2} \left(\alpha^{-1}\rhos{P^0}+\alpha^{-2}\rhos{P^0}^2\right)\norm{\theta_0}^2
\end{align}

\end{theorem}
\begin{comment}
Consider the LSA with $b_t=0$, and $P^0$ has all the mass concentrated on $A_{P^0}=\begin{bmatrix} \lambda_{\min} &0\\ 0& \lambda_{\max}\end{bmatrix}$, for some $\lambda_{\max}>\lambda_{\min}>0$. Note that in this example $\ts=0$.
By choosing $\alpha<\frac2{\lambda_{\max}}$, in this case it is straightforward to write the expression for $\eh_t$ explicitly as below:
\begin{align*}
\eh_t&=\frac{1}{t+1}\sum_{s=0}^t e_t = \frac{1}{t+1}\sum_{s=0}^t (I-\alpha A_P)^{t-s} e_0\\
&=\frac{1}{t+1}(\alpha A)^{-1}\left(I-(I-\alpha A)^{t+1}\right)e_0
\end{align*}

@@

\item \textbf{Role of $\rhos{P}$ and $\rhod{P}$:} We can expand the MSE as \begin{align}\label{eq:exp}\EE{\norm{\eh_t}}=\frac{1}{(t+1)^2}\ip{\sum_{s=0}^t e_s,\sum_{s=0}^t e_s}.\end{align} The inner product in \eqref{eq:exp} is a summation of \emph{diagonal} terms $\EE{\ip{e_i,e_i}}$ and \emph{corss} terms of$\EE{\ip{e_i,e_j}},i\neq j$. The growth of the diagonal terms and the cross terms depends on the spectral norm of the matrices $I-\alpha A_t$ and $I-\alpha A_P$ respectively. These are given by

@@

Note that \eqref{eq:spectralrand} explicitly connects $\rhod{P}$ and $\rhos{P}$ in \Cref{def:contract} to the spectral norms of the aforementioned matrices. For the MSE to be bounded, we need the spectral norms to be less than unity, which in turn translates to the requirement that $\rhos{P}>0$ and $\rhod{P}>0$.
\item \textbf{Role of $U$:} When $A_P$ is positive definite, it is straighforward to produce an $\alpha>0$ such that $\rhos{P}>0$ and $\rhod{P}>0$. However, $A_P$ is Hurwitz does not guarantee positive definiteness (see \Cref{ex:pdas}) and a result an $\alpha>0$ such that $\rhos{P}>0$ and $\rhod{P}>0$. This poses a difficulty in applying results of \Cref{th:pdrate} in the case Hurwitz matrices. This can be remedied by a similiarity transformation using an appropriate $U$ that transforms $A_P$ to appropriate Jordon form $A_{P_U}$. \Cref{th:simtran} presents the rates for the transformed iterates   $\gamma_t=U^{-1 }\theta_t$. Note that \Cref{cor:as} is obtained from \Cref{th:simtran} by the fact that $\theta_t=U\gamma_t$.

@@

\item \textbf{$\alpha$, $\rhos{P}$ and $\rhod{P}$} only affect the constants in the rates and not the exponents $\frac{1}{t}$ for variance and $\frac{1}{t^2}$ for bias terms.
\item \textbf{Behaviour for extreme value of $\alpha$:} For smaller values of stepsize, i.e., $\alpha\approx 0$, the bias term blows up, due to the presence of $\alpha^{-1}$ term. This is due to the fact that the stepsizes determine the learning rate and for smaller stepsizes the learning rate is slower. However, in this case the noise term does not blow up, a fact that can appreciated by looking at \Cref{th:pdrate,th:simtran,cor:as} where $\alpha$ is seen to multiply the variances $\sigma^2_P$ and $\sigma^2_b$. In quantitative terms, we can see that the $\alpha^{-2}$ and $\alpha^2$ terms can each other. For larger values of $\alpha$ i.e., $\alpha\ra \alpha_{P}$ (or $\alpha_{P_U}$ as the case may be), the bounds blow up again, due to the fact that $\rhos{P}\ra 0$ in this case.

\subsection{Illustrative Examples}
\begin{example}\label{ex:pdas}
Matrices $A_1=\begin{bmatrix} 2 &1 \\-1 &0\end{bmatrix}$$A_2=\begin{bmatrix}1 & -1\\ 1 &0 \end{bmatrix}$ and $A_3=\begin{bmatrix} 1 & 10-\epsilon \\ 0 &1 \end{bmatrix}$ are not positive definite. However choosing $U_1=\begin{bmatrix} \frac{1}{\sqrt{2}}& -\frac{1}{\sqrt{2}}\\ -\frac{1}{\sqrt{2}} &\frac{1}{\sqrt{2}} \end{bmatrix}$, $U_2=\begin{bmatrix} \frac{\sqrt{2}}{1-i\sqrt{3}} & \frac{\sqrt{2}}{1+i\sqrt{3}} \\ -\frac{1}{\sqrt{2}} &\frac{1}{\sqrt{2}}\end{bmatrix}$ and $U_3=\begin{bmatrix} 1 & 0\\ 0 &\frac{1}{10} \end{bmatrix}$, we can see that matrices $J_i=U_i^{-1}A_i U_i,i=1,2,3$ are positive definite.
\end{example}
\paragraph{Remark}

\begin{example}\label{ex:geo}
Consider the one dimensional case where $A_t=\rho$ for some $0<\rho<1$, $b_t$ a bounded $i.i.d$ zero mean random variable with variance $\sigma^2$, and $\ts=0$. We know that $e_t=(1-\alpha\rho)^t e_0+\sum_{s=1}^t (1-\alpha \rho)^{t-s}\alpha b_s$ and the MSE is given by
\EE{\norm{\theta_t}^2}&=\frac{1}{(t+1)^2}\left( (\alpha\rho)^{-1}(1-\alpha\rho)^{t+1}) \theta_0+ \sum_{s=1}^t (\alpha\rho)^{-1}(1-\alpha \rho)^{t-s+1} \alpha b_s\right)^2\\
&\leq (\alpha\rho)^{-2}\left(\frac{\norm{\theta_0}^2}{t^2}+\frac{\alpha^2\sigma^2}{t}\right)
\end{example}
\paragraph{Geometric Series:} In simple terms the error dynamics is akin to sum of a geometric series with common factor $\alpha\rho$, a fact that is illustrated by the simple $1$-dimensional example in \Cref{ex:geo}.
\begin{example}
Consider distributions $P^1$, $P^2$ and $P^3$ which are concentrated on matrices $A_{P^1}=\begin{bmatrix} 10 &100\\ 0 &10\end{bmatrix}$, $A_{P^2}=\begin{bmatrix} 10 &10\\ 0 &10\end{bmatrix}$, $A_{P^3}=\begin{bmatrix} 10 &0 \\ 0 &10\end{bmatrix}$. Note that $P^1$, $P^2$ and $P^3$ are respectively AS, PD, SPD respectively. Consider the LSA for the three case of $P^1$, $P^2$ and $P^3$ with $b_t=0$ and $\ts=0$
We now list the following observations:
\begin{itemize}%[leftmargin=*, before = \leavevmode\vspace{-\baselineskip}]
\item In all the three instance, $0<\alpha<\frac2{10}$ ensures that the error $\norm{\eh_t}\ra 0$ as $t\ra\infty$. Thus the range of $\alpha$ is uniform across $P^i,i=1,2,3$.
\item $\rhos{P^1}=\rhod{P^1}<0$, and hence one cannot apply results of \Cref{th:pdrate}. However, with  $U=\begin{bmatrix}1 &0 \\ 0 &0.2-\epsilon\end{bmatrix}$ for some $\epsilon>0$ we have $P^1_U$ to be the distribution concentrated on the matrix $A_{P^1_U}=\begin{bmatrix} 10 &20 \\ 0 &10\end{bmatrix}$ and one can apply the results of \Cref{thm:simtran,cor:as}. Notice that the condition number $\norm{U}{U^-1}\approx 5$.

@@

We plot $L_t$ for the three cases for $t=1,\ldots,100$ and we can see that the condition number $\norm{U}{U^-1}\approx 5$, shows up in the AS case ($P^1$).
\end{example}
\subsection{Related Work}
LSA with CS-PR was considered in \cite{polyak-judisky} under the additive noise assumption, wherein $A_t=A$. Recent work by \cite{bach} considers a stochastic gradient descent (SGD) algorithm for the problem of linear prediction given in \Cref{ex:linreg}.
\begin{example}\label{ex:linreg}
Let $(x_t,y_t)\in \R^d\times \R,\,t\geq 1$ be $i.i.d$ such that $\EE{\norm{x_t}^2}$ and $\EE{y_t^2}$ are finite. Further, we assume bounded data i.e., $\norm{(x_t,y_t)}^2\leq B$ for some positive real number $B>0$. Here, $x_t,\,t\geq 1$ are the features and $y_t$ are real labels and the linear regression problem constitutes minimizing the quadratic loss function $f(\theta)=\EE{\left(\ip{x_t,\theta}-y_t\right)^2}$. SGD with CS-PR to minimize $f(\theta)$ can be given as below:
\begin{subequations}\label{eq:linreg}
\begin{align}
\theta_t&=\theta_{t-1}-\alpha x_t\left(\ip{x_t,\theta_{t-1}} -y_t\right)\\
\thh_t&=\frac{1}{t+1}\sum_{s=0}^t \theta_t
\end{subequations}
Here $\nabla f(\theta_{t-1})=x_t\left(\ip{x_t,\theta_{t-1} -y_t}\right)$ is the \emph{stochastic gradient} of the objective $f(\theta)$. It follows from the $i.i.d$ assumption on $(x_t,y_t)$ that $A_t=x_t x_t^\top$ and $b_t=y_t x_t$ are also $i.i.d$, and with this identification one can see that \eqref{linreg} is of the form in \eqref{conststep}. In this case $A_P$ is SPD.
In the linear regression setting, the error of a parameter $\theta$ is measured as the difference between the loss at the said parameter and the loss at the optimum, which has the following form
\begin{align}\label{eq:mselinreg}
f(\theta)-f(\ts)=(\theta-\ts)^\top A_P (\theta-\ts)=\EE{\norm{\theta-\ts}^2_{A_P}}
\end{example}
In this setting, \cite{bach} show that the variance term in the MSE is at most $O(\frac{C}{t})$, where $C$ does not depend on the condition number. This ``remarkable'' result is due to the following facts: $(i)$ $A_P$ in this case is symmetric positive definite, $(ii)$ the MSE in \eqref{eq:mselinreg} is measured with respect to $A_P$, $(iii)$ the noise $\epsilon_t=y_t-\ip{\ts,x_{t-1}}$ is structured i.e., $\EE{\epsilon_t^2 x_t x_t^\top}\leq R A_P$ for some constant $R>0$.
At the same time, the bias term decays at a rate $\frac{C'}{t^2}$, where the constant $C'$ contains the factor $(\theta_0-\ts)^\top A^{-1}_P (\theta_0-\ts)$ and is dependent on the condition number.

>>>>

\section{Main Results}\label{sec:mainresults}
In this section, we derive instance dependent bounds  which is valid for a fixed $P$ and in the \Cref{sec:uniform}, address the question of deriving uniform bounds $\forall\,P\in \P$, where $\P$ is a class of distributions over $\C^{\dcd}$. Here, we only present the results followed by a discussion and the detailed proofs can be found in \Cref{sec:appendix}.
\begin{lemma}\label{lm:pd}
Let $P$ be positive definite, then there exists an $\alpha_P>0$ and $U\in \gln$ such that $\rhod{P_U}>0$ and $\rhos{P_U}>0,~\forall \alpha \in (0,\alpha_P)$.
\end{lemma}
\begin{lemma}\label{lm:simtran}[Similarity Transformation]
Let $P$ be Hurwitz, then there exists a $U\in \gln$ such that $P_U$ is positive definite.
\end{lemma}
\begin{comment}
\begin{proof}
\begin{align*}
\rhos{P}&=\inf_{x:\norm{x}=1}x^* (A_P+A_P^*)x -\alpha x^*\EE{A_t^* A_t} x\\
&=\inf_{x:\norm{x}=1}x^* (A_P+A_P^*)x -\alpha x^* A^*_P A_P -\alpha x^* \EE{M_t^* M_t} x\\
&\geq \lambda_{\min}(A^*_P+A_P)-\alpha \norm{A_P}^2-\sigma^2_P
\end{align*}
The proof is complete by choosing $\alpha_P<\frac{\lambda_{\min}(A^*_P+A_P)}{\norm{A_P}^2+\sigma^2_P}$
\end{proof}
\end{comment}

\begin{comment}
\begin{theorem}\label{th:pdrate}
Let $P^A$ in \Cref{assmp:lsa}-~\ref{dist} be a positive definite distribution over $\C^{\dcd}$. Then
\begin{align*}
\EE{\norm{\thh_t-\ts}^2}
\leq
\left(1+\frac2{\alpha\rhod{P}}\right)\frac{1}{\alpha \rhos{P}}
\left(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} \right)\,.
%\left(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\theta_0-\ts})}{t+1} \right)\,.
\end{align*}

\end{theorem}
\begin{theorem}\label{thm:simtran}[Change of Basis]
Let $P$ in \Cref{assmp:lsa}-~\ref{dist} be AS. Define $\gamma_t\eqdef U^{-1}\theta_t,\,,\gamma^*\eqdef U^{-1}\ts$, and suppose if $P_U$ is positive definite, then
\begin{align*}
\EE{\norm{\gh_t-\gamma^*}^2}
\leq
\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\norm{U^{-1}}^2}{\alpha \rhos{P_U}}
\left(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} \right)\,.
%\left(\frac{\norm{\gamma_0-\gamma^*}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\gamma_0-\gamma^*}}{t+1} \right)\,.
%\left(\frac{\norm{\gamma_0-\gamma^*}^2}{(t+1)^2}+ \frac{\alpha^2{\sigma}_1^2+ {\sigma}_2^2 (\alpha^2+\alpha \norm{\gamma_0-\gamma^*})}{t+1} \right)\,,
\end{align*}
where $\gh_t=\frac{1}{t+1}\sum_{s=0}^t \gamma_s$.
\end{theorem}
\end{comment}

\begin{theorem}\label{th:rate}
Let $P$ be Hurwitz, then there exists $U\in\gln$ and $\alpha_{P_U}>0$ such that
\begin{align*}
\EE{\norm{\thh_t-\ts}^2}
\leq
\left(1+\frac2{\alpha\rhod{P_U}}\right)\frac{\cond{U}^2}{\alpha \rhos{P_U}}
\left(\frac{\norm{\theta_0-\ts}^2}{(t+1)^2}+ \frac{\alpha^2(\sigma_P^2\norm{\ts}^2+\sigma_b^2)+\alpha (\sigma_P^2\norm{\ts})\norm{\theta_0-\ts}}{t+1} \right)\,.
\end{align*}
\end{theorem}

\begin{theorem}[Lower Bound]
There exists a $P$ Hurwitz, $\alpha_P>0$, such that $\rhos{P}>0$ and $\rhod{P}>0,\,\forall \alpha\in (0,\alpha_P)$ and
\begin{align*}
\EE{\norm{\eh_t}^2} &\geq \frac{1}{(t+1)^2}(\alpha)^{-2}(\rhod{P}\rhos{P})^{-1}\Big( \big(1-(1-\alpha \rhos{P})^t\big) \norm{\theta_0-\ts}^2 \\&+ \sum_{s=1}^t \big(1-(1-\alpha \rhos{P})^{(t-s)}\big) \alpha^2 \sigma^2_b \Big)
\end{align*}

\begin{comment}
\begin{align}
\frac{1}{(t+1)^2}(\alpha \lambda_{\min})^{-2}(1-\alpha \lambda_{\min}^t) \theta^2_0(1)\leq \norm{\eh_t}^2\leq
\frac{1}{(t+1)^2} \left(\alpha^{-1}\rhos{P^0}+\alpha^{-2}\rhos{P^0}^2\right)\norm{\theta_0}^2
\end{align}
\end{comment}
\end{theorem}

Consider the LSA with $b_t=0$, and $P^0$ has all the mass concentrated on $A_{P^0}=\begin{bmatrix} \lambda_{\min} &0\\ 0& \lambda_{\max}\end{bmatrix}$, for some $\lambda_{\max}>\lambda_{\min}>0$. Note that in this example $\ts=0$.
By choosing $\alpha<\frac2{\lambda_{\max}}$, in this case it is straightforward to write the expression for $\eh_t$ explicitly as below:
\begin{align*}
\eh_t&=\frac{1}{t+1}\sum_{s=0}^t e_t = \frac{1}{t+1}\sum_{s=0}^t (I-\alpha A_P)^{t-s} e_0\\
&=\frac{1}{t+1}(\alpha A)^{-1}\left(I-(I-\alpha A)^{t+1}\right)e_0
\end{align*}

@@

\item \textbf{Role of $U$:} When the given problem $P$ is positive definite, it is straighforward to produce an $\alpha_P>0$ such that $\rhos{P}>0$ and $\rhod{P}>0,\,\forall \alpha\in(0,\alpha_P)$. However, the problem $P$ being Hurwitz does not guarantee positive definiteness (see \Cref{ex:pdas}). This can be remedied by a similiarity transformation using an appropriate $U$ that transforms a Hurwitz problem $P$ to a positive definite problem $P_U$. To see this, let $\gamma_t\eqdef U^{-1 }\theta_t$ be a basis transformation, and let us re-write \eqref{conststep} in $\gamma_t$ by pre-multiplying \eqref{constste[} by $U^{-1}$ as shown below:
U^{-1}\theta_t&=U^{-1}\theta_{t-1}+\alpha(U^{-1} b_t- U^{-1}A_t U U^{-1}\theta_{t-1}),\\
\gamma_t&=\gamma_{t-1}+\alpha(U^{-1} b_t- J_t\gamma_{t-1}),
This completes the transformation of a Hurwirz problem $P$ to a positive definite problem $P_U$. One can revert to $\theta_t$ by using $\theta_t=U\gamma_t$, and this explains the reason for the condition number $\cond{U}$ in \Cref{th:rate}. We would like to add that when the problem $P$ is positive definite, we can let $U=\I$ in \Cref{th:rate}. We would like to mention that such similarity transformations have been in the past for analyzing LSA algorithms in RL \cite{lihong} (see \Cref{sec:related} for more details).
\item \textbf{Role of $\rhos{P}$ and $\rhod{P}$:} When $P$ is positive definite, we can expand the MSE as \begin{align}\label{eq:exp}\EE{\norm{\eh_t}}=\frac{1}{(t+1)^2}\ip{\sum_{s=0}^t e_s,\sum_{s=0}^t e_s}.\end{align} The inner product in \eqref{eq:exp} is a summation of \emph{diagonal} terms $\EE{\ip{e_i,e_i}}$ and \emph{cross} terms of$\EE{\ip{e_i,e_j}},i\neq j$. The growth of the diagonal terms and the cross terms depends on the spectral norm of the random matrix $H_t=I-\alpha A_t$ and the deterministic matrix $H_P=I-\alpha A_P$ respectively. These are given by

@@

Note that \eqref{eq:spectralrand} explicitly connects $\rhod{P}$ and $\rhos{P}$ in \Cref{def:dist} to the spectral norms matrices $H_t$ and $H_P$. For the MSE to be bounded, we need the spectral norms to be less than unity, which in turn translates to the requirement that $\rhos{P}>0$ and $\rhod{P}>0$. If $P$ is Hurwitz, we can argue on similar lines by first transforming it into a positive definite problem $P_U$ and replacing $\rhos{P}$ and $\rhod{P}$ by $\rhos{P_U}$ and $\rhod{P_U}$, and introducing $\cond{U}$ to account for the forward ($\gamma=U^{-1}\theta$) and reverse ($\theta=U\gamma$) transformations using $U^{-1}$ and $U$ respectively.
\item \textbf{Constants} {$\alpha$, $\rhos{P}$ and $\rhod{P}$} do not affect the exponents $\frac{1}{t}$ for variance and $\frac{1}{t^2}$ for bias terms.

@@

\item \textbf{Behaviour for extreme value of $\alpha$:} For smaller values of stepsize, i.e., $\alpha\approx 0$, the bias term blows up, due to the presence of $\alpha^{-1}$ term. This is due to the fact that the stepsizes determine the learning rate and for smaller stepsizes the learning rate is slower. However, in this case the noise term does not blow up, a fact that can appreciated by looking at \Cref{th:rate} where $\alpha$ is seen to multiply the variances $\sigma^2_P$ and $\sigma^2_b$. In quantitative terms, we can see that the $\alpha^{-2}$ and $\alpha^2$ terms can each other. For larger values of $\alpha$ i.e., $\alpha\ra \alpha_{P}$ (or $\alpha_{P_U}$ as the case may be), the bounds blow up again, due to the fact that $\rhos{P}\ra 0$ in this case.
\item \textbf{Lower Bound} in \Cref{th:lb} shows that

@@

\input{illustrate}
